### 2014-06-03 stat246/hw7.tex
s7

To pare down the right side, first note that the (non-empirical) expectation is given by
\[E(f(X_i,X_{N_i})|X_{N_i}=x_{N_i})=\sum_{x_i}f(x_i,x_{N_i})P(x_i|x_{N_i})=\sum_{x_i}f(x_i, x_{N_i})\frac{\exp\left(\beta f(x_i,x_{N_i})\right)}{\sum_{x_i'}\exp\left(\beta f(x_i',x_{N_i})\right)}\]
Conveniently, this is the same as the inner summand of the RHS, so we can write the RHS as
\[\sum_{i=1}^d\sum_{n=1}^NE(f(X_i^{(n)},X_{N_i}^{(n)})|X_{N_i})=\sum_{i=1}^d\sum_{x_{N_i}}E(f(X_i,X_{N_i})|X_{N_i}=x_{N_i})\left|\{n:X_{N_i}=x_{N_i}\}\right|\]
Then, trimming away the leftover counts (which are attached to the same terms on both sides) gives us the desired equality.
\ssn{d}
In the same vein as (b), the log-pseudolikelihood is
\[\sum_{n=1}^N\sum_{i=1}^d\left(\sum_{C\in C(i)}\beta_C\psi_{C}(X_{C})-\log\left(\sum_{X_i}\exp\left(\sum_{C\in C(i)}\beta_C\psi_C(X_C)\right)\right)\right)\]
Fix a clique $C$ and differentiate wrt $\beta_C$ to obtain
\[\sum_{n=1}^N\sum_{i\in C}\left(\psi_C(X_C)-\frac{\sum_{x_i}\psi_C(X_C)\exp\left(\sum_{C'\in C(i)}\beta_{C'}\psi_{C'}(X_{C'})\right)}{\sum_{x_i}\exp\left(\sum_{C'\in C(i)}\beta_{C'}\psi_{C'}(X_{C'})\right)}\right)=0\]
Then, going through the same steps as in (c), we see that the result is the same, but with $f(x_i,X_{N_i})=\psi_C(X_C)$ for every clique $C$ that contains $i$.
\subsection*{2}
Performing the multiplication $P_nP=P_{n+1}$ to get a recurrence, we have
\[\openm1-a_n&a_n\\b_n&1-b_n\closem\openm1-a&a\\b&1-b\closem=\openm1-(a+a_n(1-a-b))&a+a_n(1-a-b)\\b+b_n(1-a-b)&1-(b+b_n(1-a-b))\closem\]
so the recurrence is $a_{n+1}=a+a_n(1-a-b)$ and similar for $b_{n+1}$. To find the steady-state solution, set $a_n=a_{n+1}=a^*$ to obtain $a^*=a+a^*(1-a-b)$ or $a^*=\frac{a}{a+b}$ (and similarly $b^*=\frac{b}{a+b}$). This gives us the limit of $P^n$ as 
\[\openm\frac{b}{a+b}&\frac{a}{a+b}\\\frac{b}{a+b}&\frac{a}{a+b}\closem\]
To find closed-form formulas for $a_n$ and $b_n$, note that $a_{n+1}-a^*=(a_n-a^*)(1-a-b)$, or $a_n-a^*=(1-a-b)^{n-1}(a-a^*)$. The same goes for $b_n$.
\subsection*{3}
\ssn{a}
\vspace{130pt}
\ssn{b}
To construct $Q$ (for $n\geq3$), start with a tridiagonal matrix with $\beta$ on the diagonal and $-\alpha$ on the off-diagonals, then place $-\alpha$ at the upper-right and lower-left corners. Then, taking $x^TQx$ results in 
\[\sum_{i=1}^nx_i(-\alpha x_{i-1}+\beta x_i-\alpha x_{i+1})\]
which simplifies to $\sum_{i=1}^n-2\alpha x_ix_{i+1}+\beta x_i^2$ when we combine the like terms across summands. Multiplying by $-\frac{1}{2}$ gives us the term in the exponent.
\ssn{c}
The $l$th entry of $Qv_k$ is $\frac{\beta}{\sqrt{n}}\exp(2\pi i\frac{l}{n}k)-\frac{\alpha}{\sqrt{n}}\left(\exp(2\pi i\frac{l-1}{n}k)+\exp(2\pi i\frac{l+1}{n}k)\right)$. This is a real multiple of $\exp(2\pi i\frac{l}{n}k)$ since adding the two $\alpha$-multiplied vectors results in something with the same argument. For $k$ fixed, the multiplier is the same for every $l$, and its value is $\frac{\beta-2\alpha\cos(2\pi k/n)}{\sqrt{n}}$. Thus, the $v_k$ are eigenvectors. Further, since we know that $\beta>2\alpha$, all the eigenvalues are positive, which implies that $Q$ is strictly positive definite.
\ssn{d}
The conditional variance of $x_i$ conditioned upon all the others is $\frac{1}{Q_{ii}}$, and the conditional mean is $-\frac{1}{Q_{ii}}Q_i^Tx$, with the $i$th coordinate being excluded from the inner product.
\ssn{e}
Plots of the trajectories for $\ep=0.01$:

>>>>

To pare down the right side, first note that the (non-empirical) expectation is given by
\[E(f(X_i,X_{N_i})|X_{N_i}=x_{N_i})=\sum_{x_i}f(x_i,x_{N_i})P(x_i|x_{N_i})=\sum_{x_i}f(x_i, x_{N_i})\frac{\exp\left(\beta f(x_i,x_{N_i})\right)}{\sum_{x_i'}\exp\left(\beta f(x_i',x_{N_i})\right)}\]
Conveniently, this is the same as the inner summand of the RHS, so we can write the RHS as
\[\sum_{i=1}^d\sum_{n=1}^NE(f(X_i^{(n)},X_{N_i}^{(n)})|X_{N_i})=\sum_{i=1}^d\sum_{x_{N_i}}E(f(X_i,X_{N_i})|X_{N_i}=x_{N_i})\left|\{n:X_{N_i}=x_{N_i}\}\right|\]
Thus we have the desired equality (after the Monday correction).
\ssn{d}
In the same vein as (b), the log-pseudolikelihood is
\[\sum_{n=1}^N\sum_{i=1}^d\left(\sum_{C\in C(i)}\beta_C\psi_{C}(X_{C})-\log\left(\sum_{X_i}\exp\left(\sum_{C\in C(i)}\beta_C\psi_C(X_C)\right)\right)\right)\]
Fix a clique $C$ and differentiate wrt $\beta_C$ to obtain
\[\sum_{n=1}^N\sum_{i\in C}\left(\psi_C(X_C)-\frac{\sum_{x_i}\psi_C(X_C)\exp\left(\sum_{C'\in C(i)}\beta_{C'}\psi_{C'}(X_{C'})\right)}{\sum_{x_i}\exp\left(\sum_{C'\in C(i)}\beta_{C'}\psi_{C'}(X_{C'})\right)}\right)=0\]
Then, going through the same steps as in (c), we see that the result is the same, but with $f(x_i,X_{N_i})=\psi_C(X_C)$ for every clique $C$ that contains $i$.
\subsection*{2}
Performing the multiplication $P_nP=P_{n+1}$ to get a recurrence, we have
\[\openm1-a_n&a_n\\b_n&1-b_n\closem\openm1-a&a\\b&1-b\closem=\openm1-(a+a_n(1-a-b))&a+a_n(1-a-b)\\b+b_n(1-a-b)&1-(b+b_n(1-a-b))\closem\]
so the recurrence is $a_{n+1}=a+a_n(1-a-b)$ and similar for $b_{n+1}$. To find the steady-state solution, set $a_n=a_{n+1}=a^*$ to obtain $a^*=a+a^*(1-a-b)$ or $a^*=\frac{a}{a+b}$ (and similarly $b^*=\frac{b}{a+b}$). This gives us the limit of $P^n$ as 
\[\openm\frac{b}{a+b}&\frac{a}{a+b}\\\frac{b}{a+b}&\frac{a}{a+b}\closem\]
To find closed-form formulas for $a_n$ and $b_n$, note that $a_{n+1}-a^*=(a_n-a^*)(1-a-b)$, or $a_n-a^*=(1-a-b)^{n-1}(a-a^*)$. The same goes for $b_n$.
\subsection*{3}
\ssn{a}
\vspace{130pt}
\ssn{b}
To construct $Q$ (for $n\geq3$), start with a tridiagonal matrix with $\beta$ on the diagonal and $-\alpha$ on the off-diagonals, then place $-\alpha$ at the upper-right and lower-left corners. Then, taking $x^TQx$ results in 
\[\sum_{i=1}^nx_i(-\alpha x_{i-1}+\beta x_i-\alpha x_{i+1})\]
which simplifies to $\sum_{i=1}^n-2\alpha x_ix_{i+1}+\beta x_i^2$ when we combine the like terms across summands. Multiplying by $-\frac{1}{2}$ gives us the term in the exponent.
\ssn{c}
The $l$th entry of $Qv_k$ is $\frac{\beta}{\sqrt{n}}\exp(2\pi i\frac{l}{n}k)-\frac{\alpha}{\sqrt{n}}\left(\exp(2\pi i\frac{l-1}{n}k)+\exp(2\pi i\frac{l+1}{n}k)\right)$. This is a real multiple of $\exp(2\pi i\frac{l}{n}k)$ since adding the two $\alpha$-multiplied vectors results in something with the same argument. For $k$ fixed, the multiplier is the same for every $l$, and its value is $\frac{\beta-2\alpha\cos(2\pi k/n)}{\sqrt{n}}$. Thus, the $v_k$ are eigenvectors. Further, since we know that $\beta>2\alpha$, all the eigenvalues are positive, which implies that $Q$ is strictly positive definite.
\ssn{d}
The conditional variance of $x_i$ conditioned upon all the others is $\frac{1}{Q_{ii}}$, and the conditional mean is $-\frac{1}{Q_{ii}}Q_i^Tx$, with the $i$th coordinate being excluded from the inner product.
\ssn{e}
Plots of the trajectories for $\ep=0.01$:
